---
title: "Prototype measurement framework: Prescribing of Breast cancer medicines"
author: "Impact team"
date: "December 2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Prevents data loading in with numbers in scientific format
options(scipen=999)

# Load packages
library(tidyverse)
library(reactable)
library(plotly)
library(downloadthis)

# Identify project file
here::i_am("Prototype_BC_Prescribing.Rproj")

# Load in functions
source(here::here("scripts/rmd_functions.R"))

```

```{r load_data, eval = TRUE}

# Load in TA table
meds_ta_df <- readr::read_csv(here::here("data/clean/meds_ta_df.csv"), 
                                col_select = 1:3)

# Load in national level data
nat_meds_df <- readr::read_csv(here::here("data/clean/bc_meds_nat_2022_10.csv"))

```

# Medicine comparison

```{r all_meds_plot, out.width = "100%"}

chart_title <- "<b>Prescribing of breast cancer medicines</b>"

nat_meds_df %>% 
    filter(treatment_name != "pembrolizumab") %>% 
    group_by(date, treatment_name) %>% 
    summarise(numerator = sum(numerator),
              denominator = first(denominator),
              dd_per_100000 = numerator/(denominator/100000)) %>% 
    ungroup() %>% 
    plot_ly(x = ~date,
            y = ~dd_per_100000,
            color = ~treatment_name) %>% 
    add_trace(type = "scatter",
              mode = "lines+markers") %>% 
    layout(title = list(text = stringr::str_wrap(chart_title, width = 95),
                        font = list(size = 12),
                        x = 0.05,
                        y = 0.96,
                        yanchor = "top"),
           yaxis = list(title = list(text = "Daily doses per 100,000 population",
                                     standoff = 5),
                        rangemode = "tozero"),
           xaxis = list(title = FALSE),
           hovermode = "x unified",
           margin = list(t = 65)) %>% 
    config(displaylogo = FALSE,
           modeBarButtonsToRemove = c("zoom", "pan", "select", "lasso"))

```

# Individual medicines

## Abemaciclib {.tabset}

### Observed usage

```{r abemaciclib_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "abemaciclib", 
                plot_title = "<b>Observed vs expected usage of abemaciclib</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "abemaciclib")

```

### Underlying data

```{r abemaciclib_data}

nat_meds_df %>% 
    filter(treatment_name == "abemaciclib") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))

```

### Relevant guidance

```{r abemaciclib_TAs}
create_med_table(meds_ta_df, "abemaciclib")

```

### Therapeutic indications  

#### Early Breast Cancer

Abemiciclib in combination with endocrine therapy is indicated for the adjuvant treatment of adult patients with hormone receptor (HR)-positive, human epidermal growth factor receptor 2 (HER2)-negative, node-positive early breast cancer at high risk of recurrence.  

In pre- or perimenopausal women, aromatase inhibitor endocrine therapy should be combined with a luteinising hormone-releasing hormone (LHRH) agonist.
&nbsp;  


#### Advanced or Metastatic Breast Cancer 

Abemiciclib is indicated for the treatment of women with hormone receptor (HR)-positive, human epidermal growth factor receptor 2 (HER2)-negative locally advanced or metastatic breast cancer in combination with an aromatase inhibitor or fulvestrant as initial endocrine-based therapy, or in women who have received prior endocrine therapy.
In pre- or perimenopausal women, the endocrine therapy should be combined with a LHRH agonist.

&nbsp;  

### Eligible population

Around 4,000 people are expected to receive treatment with abemaciclib in England annually.  

This has been taken from the Resource impact template developed for the NICE technology appraisal 'Abemaciclib with endocrine therapy for adjuvant treatment of hormone receptor-positive, HER2-negative, node-positive early breast cancer at high risk of recurrence', published July 2022.
&nbsp;  

```{r abemaciclib_pop }

abemaciclib_estimate <- read_csv(here::here("data/Abemaciclib_estimate.csv"))

reactable(abemaciclib_estimate,
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))

```
&nbsp;  

## Neratinib {.tabset}

### Observed usage

```{r neratinib_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "neratinib", 
                plot_title = "<b>National quarterly usage of neratinib</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "neratinib")

```
### Underlying data
```{r neratinib_data}

nat_meds_df %>% 
    filter(treatment_name == "neratinib") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r neratinib_TAs}
create_med_table(meds_ta_df, "neratinib")

```

&nbsp;  

## Palbociclib {.tabset}
### Observed usage
```{r palbociclib_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "palbociclib", 
                plot_title = "<b>National quarterly usage of palbociclib</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "palbociclib")

```
### Underlying data
```{r palbociclib_data}

nat_meds_df %>% 
    filter(treatment_name == "palbociclib") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r palbociclib_TAs}
create_med_table(meds_ta_df, "palbociclib")

```

&nbsp;  

## Pembrolizumab {.tabset}
### Observed usage
```{r pembrolizumab_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "pembrolizumab", 
                plot_title = "<b>National quarterly usage of pembrolizumab</b>", 
                ylabel = "Milligrams") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "pembrolizumab")

```
### Underlying data
```{r pembrolizumab_data}

nat_meds_df %>% 
    filter(treatment_name == "pembrolizumab") %>% 
    group_by(date, treatment_name) %>% 
    summarise("mgs" = sum(numerator),
              "mgs per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r pembrolizumab_TAs}
create_med_table(meds_ta_df, "pembrolizumab")

```

&nbsp;  

## Ribociclib {.tabset}
### Observed usage
```{r ribociclib_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "ribociclib", 
                plot_title = "<b>National quarterly usage of ribociclib</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "ribociclib")

```

### Underlying data
```{r ribociclib_data}

nat_meds_df %>% 
    filter(treatment_name == "ribociclib") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r ribociclib_TAs}
create_med_table(meds_ta_df, "ribociclib")

```

&nbsp;  

## Trastuzumab deruxtecan {.tabset}
### Observed usage
```{r trast_der_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "trastuzumab deruxtecan", 
                plot_title = "<b>National quarterly usage of trastuzumab deruxtecan</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "trastuzumab deruxtecan")

```
### Underlying data
```{r trast_der_data}

nat_meds_df %>% 
    filter(treatment_name == "trastuzumab deruxtecan") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance
```{r trastuzumab_der_TAs}
create_med_table(meds_ta_df, "trastuzumab deruxtecan")

```

&nbsp;  

## Trastuzumab emtansine {.tabset}
### Observed usage

```{r trast_em_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "trastuzumab emtansine", 
                plot_title = "<b>National quarterly usage of trastuzumab emtansine</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "trastuzumab emtansine")

```
### Underlying data
```{r trast_em_data}

nat_meds_df %>% 
    filter(treatment_name == "trastuzumab emtansine") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r trastuzumab_emt_TAs}
create_med_table(meds_ta_df, "trastuzumab emtansine")

```

&nbsp;  

## Tucatinib {.tabset}
### Observed usage
```{r tucatinib_usage_plot, out.width = "100%"}

single_med_plot(nat_meds_df, 
                medicine = "tucatinib", 
                plot_title = "<b>National quarterly usage of tucatinib</b>", 
                ylabel = "People") %>% 
        add_TA_lines(meds_ta_df, nat_meds_df, "tucatinib")

```
### Underlying data
```{r tucatinib_data}

nat_meds_df %>% 
    filter(treatment_name == "tucatinib") %>% 
    group_by(date, treatment_name) %>% 
    summarise("Daily doses" = sum(numerator),
              "Doses per 100,000 population" = round(sum(numerator)/(first(denominator)/100000), 2),
              # People, assuming quarter is 91 days
              "Estimated people" = round(numerator/91,0)) %>% 
    ungroup() %>% 
    arrange(desc(date)) %>% 
    rename("Date" = date, "Medicine" = treatment_name) %>% 
    reactable(
        filterable = FALSE,
        searchable = FALSE,
        highlight = TRUE,
        style = list(fontSize = "14px"),
        defaultColDef = colDef(
            format = colFormat(separators = TRUE)))
```
### Relevant guidance

```{r tucatinib}
create_med_table(meds_ta_df, "tucatinib")

```